### 2023-09-26 doc/_bayes_priors.qmd
finish prior definition section

One aspect commonly associated---in both positive and negative ways---to
Bayesian analysis is the definition of prior probabilities. On the one hand, the
naive textbook version of Bayesian analysis defines the prior as encoding the
information we have about the inference in question before acquiring any data.
This is the "ideal" use of priors that, whenever possible, should be used. On
the other hand, for most practitioners of Bayesian statistics in the age of big
data, the definition of prior becomes a tool to ensure the convergence of
sampling algorithms such as MCMC [@gelman2017]. However, for our particular
problem, although we deal with large amounts of data (inferences can be made for
> 10K barcodes over multiple time points, resulting in > 100K parameters), each
barcode has very little data, as they are measured only once per time point over
< 10 times. Furthermore, understanding the sources of noise having to do with
culturing conditions, DNA extraction, library preparation, and so on and
encoding them into reasonable prior distributions becomes incredibly
challenging.

@@

With this in mind, we leave it to the reader to judge the selection of priors.
Furthermore, the present package `BayesFitness.jl` is written so that users can
experiment with different prior selection criteria that fit their needs. We
strongly advocate that statistics should not be done in a black-box fit-all tool
mindset but rather as a formal way to encode the assumptions behind the
analysis, subject to constructive criticism.

With this philosophical baggage behind us, let us now focus on how the priors
used for this work were selected.

@@

Furthermore, the `naive_prior` function defines the mean of the variance
parameter as the standard deviation of the log frequency ratios for the neutral
lineages, i.e.,
$$
\mu_{\sigma_t} = \sqrt{
    \operatorname{Var}\left(\frac{r_{t+1}^{(n)}}{r_t^{(n)}}\right)
}.
$${#eq-prior_meanfit_var_data}
This same value was utilized for the

>>>>

One aspect commonly associated---in both positive and negative ways---to
Bayesian analysis is the definition of prior probabilities. On the one hand, the
naive textbook version of Bayesian analysis defines the prior as encoding the
information we have about the inference in question before acquiring any data.
This is the "ideal" use of priors that, whenever possible, should be
implemented. On the other hand, for most practitioners of Bayesian statistics in
the age of big data, the definition of prior becomes a tool to ensure the
convergence of sampling algorithms such as MCMC [@gelman2017]. However, for our
particular problem, although we deal with large amounts of data (inferences can
be made for > 10K barcodes over multiple time points, resulting in > 100K
parameters), each barcode has very little data, as they are measured only once
per time point over < 10 growth-dilution cycles. Furthermore, it is incredibly
challenging to understand the noise sources related to culturing conditions, DNA
extraction, library preparation, etc., and encode them into reasonable
prior distributions.

@@

With this in mind, we leave it to the reader to judge the selection of priors.
Furthermore, the software package associated with this work, `BayesFitness.jl`,
is written so that users can experiment with different prior selection criteria
that fit their needs. We strongly advocate that statistics should not be done in
a black-box fit-all tool mindset but rather as a formal way to encode the
assumptions behind the analysis, subject to constructive criticism. With this
philosophical baggage behind us, let us now focus on how the priors used for
this work were selected.

@@

Furthermore, the `naive_prior` function defines the mean of the variance
parameter as the standard deviation of the log frequency ratios for the neutral
lineages, i.e.,
$$
\mu_{\sigma_t} = \sqrt{
    \operatorname{Var}\left(\frac{r_{t+1}^{(n)}}{r_t^{(n)}}\right)
},
$${#eq-prior_meanfit_var_data}
where $\operatorname{Var}$ is the sample variance. This same value was utilized
for the mean of the non-neutral barcode variance $\mu_{\sigma^{(m)}}$. While we
assigne the corresponding variances to be $\sigma_{\sigma_t} =
\sigma_{\sigma^{(m)}} = 1$.
